---
title: "Macro View"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: NULL
    logo: inst/imgs/logo.png
    social: menu
---


<style>                     
.navbar {
  background-color:#282f6b;
}
</style> 


```{r setup, include=FALSE}

# Setup and load packages

knitr::opts_knit$set(progress = TRUE)

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  "rmarkdown",
  "highcharter",
  "tidyverse",
  "flexdashboard",
  "htmlwidgets",
  "kableExtra",
  "gt",
  "scales",
  "colorspace",
  "magrittr",
  "rlist"
  )

# Set the default language of date in R
Sys.setlocale("LC_TIME", "English")

```

```{r data, include=FALSE}

# Load RDATA files

c("econ_activity", "fiscal", "inflation", "international", "monetary", "labor") %>%
  paste0("./inst/extdata/", ., ".Rdata") %>% 
  purrr::map( ~ load(.x, .GlobalEnv))

source("./R/hc_theme.R")
source("./R/ftplot.R")
source("./R/utils.R")
```

<!-- Overview section -->

Economic activity {data-navmenu="Sections" data-icon="fa-chart-line"}
=====================================

Column {data-width=375}
-----------------------------------------------------------------------


### GDP growth in 4 quarters (`r gdp_growth$date` - IBGE)

```{r section1_gdp_growth}
gdp_growth %>%
  pull(value) %>%
  valueBox(
    value = .,
    icon  = "fa-chart-line",
    color =  ifelse(. < 0, "#b22200", "#282f6b")
    )
```

### GDP at current prices in 4 quarters (`r gdp_cur_prices$date` - IBGE)

```{r section1_gdp_current_prices}
gdp_cur_prices %>%
  pull(value) %>%
  valueBox(
    value = .,
    icon  = "fa-dollar-sign",
    color = "#224f20"
    )
```

### Cielo retail index (`r format(icva$date, "%B, %Y")` - ICVA)

```{r section1_icva}
icva %>%
  pull(value) %>%
  valueBox(
    value = .,
    icon  = "fa-store",
    color =  ifelse(. < 0, "#b22200", "#282f6b")
    )
```

### Vehicle production (`r format(vehicle$date, "%B, %Y")` - ANFAVEA)

```{r section1_vehicle_production}
vehicle %>%
  pull(value) %>%
  valueBox(
    value = .,
    icon  = "fa-car",
    color = "#5f487c"
    )
```

### Installed Capacity Utilization Level (`r format(nuci$date, "%B, %Y")` - FGV)

```{r section1_nuci_fgv}
gauge(
  value  = nuci$value,
  min    = 0,
  max    = 100,
  symbol = "%",
  gaugeSectors(
    success = c(85, 100),
    warning = c(50, 85),
    danger  = c(0, 50),
    colors = c("#224f20", "#b35c1e", "#b22200")
    )
  )
```



Column {data-width=575}
-----------------------------------------------------------------------


### GDP trend {.no-title}

```{r section1_gdp_trend}
gdp_growth_sector %>%
  hchart("column", hcaes(x = date, y = value, group = sector), color = colors[1:4]) %>%
  ftplot(
    title    = "GDP growth by sectors",
    subtitle = "Accumulated growth rate in 4 quarters",
    source   = "IBGE"
    )
```

### IBC-Br trend {.no-title}

```{r section1_ibcbr_trend}
ibc_growth %>%
  hchart("line", hcaes(x = date, y = yoy, group = variable), color = colors) %>%
  ftplot(
    title    = "IBC-Br growth by region",
    subtitle = "Accumulated growth rate in 12 months",
    source   = "BCB"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    )
```


Column {data-width=500 .tabset}
-----------------------------------------------------------------------


### GDP approaches

```{r section1_gdp_approaches}
gdp_measures %>%
  gt(rowname_col = "sector") %>%
  tab_source_note(footnote_gdp) %>% 
  tab_source_note("Source: IBGE") %>%
  tab_footnote(
    footnote = "Growth rate for the quarter versus the immediately preceding quarter",
    locations = cells_column_labels(columns = vars(`Quarter over Quarter`))
    ) %>% 
  tab_footnote(
    footnote = "Quarterly growth rate (compared to the same period of the previous year)",
    locations = cells_column_labels(columns = vars(`Year over Year`))
    ) %>% 
  tab_footnote(
    footnote = "Accumulated growth rate over the year (compared to the same period of the previous year)",
    locations = cells_column_labels(columns = vars(`Year to Date`))
    ) %>% 
  tab_footnote(
    footnote = "Growth rate accumulated in 4 quarters (compared to the same period of the previous year)",
    locations = cells_column_labels(columns = vars(`Accumulated in 4 Quarters`))
    ) %>%
  data_color(
    columns = 2:5,
    colors = scales::col_numeric(
      colorspace::diverge_hcl(n = 20, palette = "Blue-Red 3"),
      reverse = TRUE,
      domain = c(-max_abs(gdp_measures[2:5]), max_abs(gdp_measures[2:5]))
      )
    ) %>%
  fmt_percent(columns = 2:5, scale_values = FALSE, decimals = 1) %>%
  cols_width(2:5 ~ px(125)) %>% 
  tab_options(
    column_labels.font.weight = "bold",
    stub.font.weight = "bold",
    footnotes.padding = 3,
    data_row.padding = 6
    )
```

### Retail (PMC)

```{r section1_pmc}
pmc %>%
  hchart("line", hcaes(x = date, y = value, group = variable), color = colors[1:4]) %>%
  ftplot(
    title    = "Retail trade index: PMC/IBGE",
    subtitle = "Monthly percentage change (compared to the same month in the previous year)",
    source   = "IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```

### Services (PMS)

```{r section1_pms}
pms %>%
  hchart("line", hcaes(x = date, y = value, group = variable), color = colors[1:2]) %>%
  ftplot(
    title    = "Revenue and volume of services index: PMS/IBGE",
    subtitle = "Monthly percentage change (compared to the same month in the previous year)",
    source   = "IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```

### Industry (PIM-PF)

```{r section1_pim}
pim %>%
  hchart("line", hcaes(x = date, y = value, group = variable), color = colors[1:4]) %>%
  ftplot(
    title    = "Industrial production: PIM-PF/IBGE",
    subtitle = "Monthly percentage change (compared to the same month in the previous year)",
    source   = "IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```


Inflation {data-navmenu="Sections" data-icon="fa-hand-holding-usd"}
=====================================

Column {data-width=350}
-----------------------------------------------------------------------


### Inflation over 12 months (`r format(ipca_yoy$date, "%B, %Y")` - IPCA/IBGE)

```{r section2_inflation_12m}
ipca_yoy %>%
  pull(value) %>%
  valueBox(
    value = .,
    icon  = "fa-chart-bar",
    color =  ifelse(. > inflation_target$value, "#b22200", "#282f6b")
    )
```

### Inflation target for (`r inflation_target$date` - CMN/BCB)

```{r section2_inflation_target}
inflation_target %>% 
  mutate(value = paste0(value, "%")) %>%
  pull(value) %>%
  valueBox(
    value = .,
    icon = "fa-bullseye",
    color = "#5f487c"
    )
```

### Diffusion index (IPCA) {.no-title}

```{r section2_diffusion_ipca}
ipca_diffusion %>%
  hchart(
    "area", hcaes(x = date, y = value, group = variable),
    color = colors[1], fillOpacity = 0.5) %>%
  ftplot(
    title    = "Diffusion index - IPCA",
    subtitle = "",
    source   = "BCB"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```


Column {data-width=550}
-----------------------------------------------------------------------


### Inflation trend {.no-title}

```{r section2_ipca_trend}
hchart(
  ipca_trend, "column", hcaes(x = date, y = `Month over Month (%)`),
       color = colors[1], name = "Month over Month", showInLegend = TRUE
  ) %>%
  hc_add_series(
    ipca_trend, "line", hcaes(x = date, y = `Year over Year (%)`),
    color = colors[2], name = "Year over Year", showInLegend = TRUE
  ) %>% 
  ftplot(
    title    = "Inflation trend (IPCA)",
    subtitle = "Monthly and accumulated variation in 12 months (%)",
    source   = "IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    )
```

### Inflation by metropolitan region {.no-title}

```{r section2_ipca_region}
ipca_region %>%
  hchart("column", hcaes(x = date, y = value, group = region), color = colors) %>%
  ftplot(
    title    = "Inflation by metropolitan region - IPCA",
    subtitle = "Monthly percentage change",
    source   = "IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 0, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    )
```


Column {data-width=450 .tabset}
-----------------------------------------------------------------------


### CPI components

```{r section2_ipca_groups}
ipca_groups %>%
  gt(rowname_col = "Group") %>%
  tab_source_note(footnote_ipca_groups) %>% 
  tab_source_note("Source: IPCA/IBGE") %>%
  tab_footnote(
    footnote = "Month over Month: monthly percentage change",
    locations = cells_column_labels(columns = vars(`MoM`))
    ) %>% 
  tab_footnote(
    footnote = "Year to Date: percentage change accumulated in the year",
    locations = cells_column_labels(columns = vars(`YTD`))
    ) %>% 
  data_color(
    columns = 2:3,
    colors = scales::col_numeric(
      colorspace::diverge_hcl(n = 20, palette = "Blue-Red 3"),
      reverse = TRUE,
      domain = c(-max_abs(ipca_groups[2:3]), max_abs(ipca_groups[2:3]))
      )
    ) %>%
  fmt_percent(columns = 2:3, scale_values = FALSE, decimals = 1) %>%
  cols_width(2:3 ~ px(130)) %>%
  cols_width(1 ~ px(180)) %>%
  tab_options(
    column_labels.font.weight = "bold",
    stub.font.weight          = "bold",
    footnotes.padding         = 3,
    data_row.padding          = 6
    )
```

### CPI cores

```{r section2_ipca_cores}
ipca_cores %>%
  hchart("line", hcaes(x = date, y = value, group = variable), color = colors[1:5]) %>%
  ftplot(
    title    = "CPI cores (IPCA)",
    subtitle = "Monthly percentage change",
    source   = "BCB"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```

### IGP trend

```{r section2_igp}
igp %>%
  hchart("line", hcaes(x = date, y = value, group = variable), color = colors[1:3]) %>%
  ftplot(
    title    = "General Price Index (IGP)",
    subtitle = "Monthly percentage change",
    source   = "FGV"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```


Labor market {data-navmenu="Sections" data-icon="fa-users"}
=====================================

Column {data-width=375}
-----------------------------------------------------------------------

### # Employed (`r big_numbers$date` - IBGE)

```{r section3_employed}
big_numbers %>% 
  pull(employed) %>% 
  valueBox(
    value = .,
    icon = "fa-users",
    color = "#282f6b"
    )
```

### # Unemployed (`r big_numbers$date` - IBGE)

```{r section3_unemployed}
big_numbers %>% 
  pull(unemployed) %>% 
  valueBox(
    value = .,
    icon = "fa-user-times",
    color = "#b22200"
    )
```

### Employment by category {.no-title}

```{r section3_employment_category}
highchart() %>%
  hc_chart(type = "item") %>% 
  hc_title(text = "Employment by category") %>%
  hc_subtitle(
    text = paste0("In millions of employees, updated until ", employment_category$date[1])
    ) %>%
  hc_credits(enabled = TRUE, text = "Source: IBGE", style = list(fontSize = "12px")) %>%
  hc_legend(labelFormat = '{name} <span style="opacity: 0.4">{y}</span>') %>%
  hc_series(
    list(
      name = "Employees",
      keys = c("name", "y", "color"),
      data = {
        employment_category %>%
          select(-date) %>%
          mutate(color = colors) %>%
          unname() %>%
          transpose(.names = NULL)
        },
      dataLabels = list(
        enabled = TRUE,
        format = "{point.name} million"
      ),
      center = c("50%", "88%"),
      size = "50%"
    )
  )
```


Column {data-width=500}
-----------------------------------------------------------------------

### Unemployment rate {.no-title}

```{r section3_unemployment}
unemployment %>%
  hchart(
    "area", hcaes(x = date, y = unemployment_rate), color = colors[1],
    name = "Unemployment rate", fillOpacity = 0.5) %>%
  ftplot(
    title    = "Unemployment rate",
    subtitle = "Quarterly data (PNADC-T)",
    source   = "IBGE"
    )
```

### Unemployment rate by gender and color {.no-title}

```{r section3_unemployment_gender_color}
unemployment_gender_color %>%
  hchart("column", hcaes(x = color, y = unemployment_rate, group = gender), color = colors[1:2]) %>%
  ftplot(
    title    = "Unemployment rate by gender and color",
    subtitle = paste0(
      "Quarterly data, updated until ", unemployment_gender_color$date[1], " (PNADC-T)"
      ),
    source   = "IBGE"
    )
```


Column {data-width=500 .tabset}
-----------------------------------------------------------------------

### Unemployment by states

```{r section3_unemployment_states}
hcmap(
  map               = "countries/br/br-all",
  data              = unemployment_states,
  joinBy            = "woe-name",
  value             = "unemployment_rate",
  name              = "Unemployment rate",
  download_map_data = TRUE,
  dataLabels        = list(
    enabled = TRUE,
    format = "{point.name}"
    ),
  mapNavigation = list(
    enabled = TRUE,
    buttonOptions = list(
      verticalAlign = "bottom"
      )
    )
  ) %>% 
  hc_title(text = "Unemployment rate by states") %>% 
  hc_subtitle(text = paste0(
      "Quarterly data, updated until ", unemployment_states$date[1], " (PNADC-T)"
      )
      ) %>%
  hc_credits(enabled = TRUE, text = "Source: IBGE", style = list(fontSize = "12px"))
```

### Occupation

```{r section3_occupation_level}
occupation_level %>%
  hchart(
    "line", hcaes(x = date, y = value, group = gender), color = colors[1:3]) %>%
  ftplot(
    title    = "Labor force occupation level",
    subtitle = "Quarterly data (PNADC-T)",
    source   = "IBGE"
    )
```

### Participation

```{r section3_participation_rate}
participation_rate %>%
  hchart(
    "line", hcaes(x = date, y = value, group = gender), color = colors[1:3]) %>%
  ftplot(
    title    = "Labor force participation rate",
    subtitle = "Quarterly data (PNADC-T)",
    source   = "IBGE"
    )
```

### Income

```{r section3_income}
income %>%
  hchart(
    "line", hcaes(x = date, y = value, group = age), color = colors) %>%
  ftplot(
    title    = "Average real income from work, by age group",
    subtitle = "Quarterly data (PNADC-T)",
    source   = "IBGE"
    ) %>%
  hc_yAxis(labels = list(format = "R$ {value:,.0f}"))
```

### Gini

```{r section3_gini}
gini %>%
  hchart(
    "bar", hcaes(x = states, y = value), color = colors[1], name = "Gini") %>%
  ftplot(
    title    = "Gini index by states",
    subtitle = paste0(
      "Quarterly data, updated until ", gini$date[1], " (PNADC-T)"
      ),
    source   = "IBGE"
    ) %>%
  hc_yAxis(labels = list(format = "{value}"))
```


Fiscal policy {data-navmenu="Sections" data-icon="fa-gavel"}
=====================================

Column {data-width=450}
-----------------------------------------------------------------------

### Central Government Primary Deficit {.no-title}

```{r section4_primary_deficit}
primary_deficit %>%
  hchart(
    "area", hcaes(x = date, y = value), color = colors[1],
    name = "Primary Deficit", fillOpacity = 0.5
    ) %>%
  ftplot(
    title    = "Central Government Primary Deficit",
    subtitle = "Accumulated in 12 months (% GDP)",
    source   = "STN and IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```

### Central Government Net Revenue and Total Expenditure {.no-title}

```{r section4_revenue_spending}
revenue_spending %>%
  hchart("line", hcaes(x = date, y = value, group = variable), color = colors[1:2]) %>%
  ftplot(
    title    = "Central Government Net Revenue and Total Expenditure",
    subtitle = "Accumulated in 12 months and deflated by the IPCA index (% GDP)",
    source   = "STN and IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE)
```


Column {data-width=450}
-----------------------------------------------------------------------

### Detailed accounts of Central Government Income and Expenditure {.no-title}

```{r section4_revenue_spending_detail}
highchart() %>%
  hc_add_series(
    type = "treemap",
    layoutAlgorithm = "squarified",
    allowDrillToNode = TRUE,
    levels = list(
      list(level = 1, 
           dataLabels = list(enabled = TRUE, style = list(fontSize = "25px")), 
           borderWidth = 8
           ),
      list(level = 2, dataLabels = list(enabled = TRUE), borderWidth = 5),
      list(level = 3, dataLabels = list(enabled = TRUE), borderWidth = 3)
    ),
    tooltip = list(valueDecimals = 2, valuePrefix = "R$ "),
    data = list(
      data = list(
        # LEVEL I
        list(
          id = unique(revenue_spending_detail$group_1)[1] %>%
            stringr::str_replace_all("[[:blank:]]", "_") %>% stringr::str_to_lower(),
          name = unique(revenue_spending_detail$group_1)[1],
          color = colors[1]
          ),
        list(
          id = unique(revenue_spending_detail$group_1)[2] %>%
            stringr::str_replace_all("[[:blank:]]", "_") %>% stringr::str_to_lower(),
          name = unique(revenue_spending_detail$group_1)[2],
          color = colors[2]
          ),
        list(
          id = unique(revenue_spending_detail$group_1)[3] %>%
            stringr::str_replace_all("[[:blank:]]", "_") %>% stringr::str_to_lower(),
          name = unique(revenue_spending_detail$group_1)[3],
          color = colors[3]
          )
        )
      ) %>%
      list.append(
        # LEVEL II
        {revenue_spending_detail %>%
            select("name" = group_2, "parent" = group_1, value) %>%
            mutate(parent = stringr::str_replace_all(parent, "[[:blank:]]", "_") %>%
                     stringr::str_to_lower()) %>%
            group_by(name, parent) %>%
            summarise(value = sum(value)) %>%
            mutate(id = stringr::str_replace_all(name, "[[:blank:]]", "_") %>%
                     stringr::str_to_lower()) %>%
            transpose()
        },
        # LEVEL III
        {revenue_spending_detail %>%
            select("name" = group_3, "parent" = group_2, value) %>%
            mutate(parent = stringr::str_replace_all(parent, "[[:blank:]]", "_") %>%
                     stringr::str_to_lower()) %>%
            transpose()
        }
      ) %>%
      flatten()
    ) %>% 
   ftplot(
    title    = "Detailed accounts of Central Government Income and Expenditure",
    subtitle = sprintf(
      "Accumulated in 12 months and deflated by the IPCA index, R$ million (%s)",
      revenue_spending_detail$date[1]
      ),
    source   = "STN and IBGE"
    )
```

### Single Account balance {.no-title}

```{r section4_single_account}
single_account %>% 
  hchart(
    "area", hcaes(x = date, y = value, group = variable),
    color = colors[4], fillOpacity = 0.5) %>%
  ftplot(
    title    = "Single Account balance",
    subtitle = "Deflated by the IPCA index, R$ million",
    source   = "BCB and IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE) %>%
  hc_yAxis(labels = list(format = "R$ {value:,.0f}"))
```


Column {data-width=450 .tabset}
-----------------------------------------------------------------------

### Debt Stock

```{r section4_debt_stock}
debt_stock %>% 
  hchart(
    "area", hcaes(x = date, y = value, group = variable),
    color = colors[2], fillOpacity = 0.5) %>%
  ftplot(
    title    = "Federal Public Debt Stock",
    subtitle = "Deflated by the IPCA index, R$ billion",
    source   = "BCB and IBGE"
    ) %>%
  hc_rangeSelector(
    selected = 4, enabled = TRUE, allButtonsEnabled = TRUE,
    inputEnabled = FALSE, dropdown = "always"
    ) %>%
  hc_navigator(enabled = TRUE) %>%
  hc_yAxis(labels = list(format = "R$ {value:,.0f}"))
```

### Government Securities Portfolio

```{r section4_gov_portfolio}
gov_portfolio %>% 
  hchart(
    "pie", hcaes(y = value, name = variable, color = colors10[1:8]),
    name = "Government Securities") %>%
  ftplot(
    title    = "Details of the National Treasury Securities portfolio",
    subtitle = sprintf("R$ billion (%s)", gov_portfolio$date_my[1]),
    source   = "BCB"
    ) %>%
  hc_tooltip(
    pointFormat =  '<b>R$ {point.value:,.0f}</b>'
  )
```



Monetary policy {data-navmenu="Sections" data-icon="fa-percentage"}
=====================================




Sobre {data-icon="fa-file-alt"}
=====================================

### Sobre {.no-title}
#### **Painel Macro**
Este projeto tem objetivo de apresentar dados macroeconômicos de forma dinâmica e interativa, contemplando alguns temas e variáveis importantes para acompanhamento de conjuntura econômica. Dividido em seis temas, o painel traz possibilidade de filtrar períodos das séries temporais, desagregar dados e selecionar variáveis nos gráficos interativos. **Explore os dados**!
**Importante**: para análises dos dados consulte os [Boletins de Conjuntura](https://gecefurg.com.br/boletins/boletim-macro/) do GECE.




#### **Informações técnicas**
O painel foi criado utilizando **R**, uma linguagem de programação livre voltada para análise estatística, através da integração do **R Markdown** com o pacote `flexdashboard`, utilizado para criação de dashboards interativas. Os dados são extraídos de diversas fontes utilizando pacotes específicos disponibilizados no [CRAN](https://cran.r-project.org/), são tratados utilizando pacotes como `tidyverse`, `zoo`, `xts` e outros, e são apresentados graficamente com o pacote `D3plusR` e em tabelas com o `kableExtra`.




#### **GECE**
Criado em agosto de 2017, o GECE tem como objetivo realizar análises de conjuntura econômica no que tange os principais indicadores macroeconômicos e ambientais. Atualmente é vinculado ao Instituto de Ciências Econômicas, Administrativas e Contábeis (ICEAC) da Universidade Federal do Rio Grande (FURG). Na área macroeconômica, as análises e boletins publicados são focados em nível de atividade, inflação, mercado de trabalho, economia internacional, política fiscal e monetária. Na área ambiental, são abordados temas como agricultura orgânica, agrotóxicos, biodiversidade, pesca, energia, florestas, unidades de conservação, saneamento e poluição.




#### **Autor**
Fernando da Silva é graduado em Economia pela FURG, escreve sobre política monetária para o GECE, tem interesses em macroeconomic research, data analysis e linguagem de programação em R. Atualmente reside em Florianópolis/SC.
Contato: <fernando@gecefurg.com.br>
LinkedIn: <https://www.linkedin.com/in/schoulten/>




#### **Dúvidas/sugestões/correções**
E-mail: <contato@gecefurg.com.br>



